---
title: 'STA 141A Fall 2022: Practice Midterm Exam 1'
author: ''
output:
  html_document
---


For this exam you are only allowed to use a pen and a three-page cheat sheet. The total number of points is 30. Do not forget to write your name and your student ID number.

Name: \hspace{2.5in} Student ID:


### Data Objects

1. (2 points) What is the result of the following R code?
  
  ```{r}
X <- matrix(c(1:6), nrow = 3, ncol = 2)
(X[X[, 2] > 4, 1])
```
<!-- The vector (2, 3) -->
  
  \vspace{60pt}

2. (2 points) Explain the differences between a matrix and a data frame.
<!-- Matrices can only contain one type of variables; 
not possible to use attach() with matrices; 
cannot select the columns of a matrix using the dollar sign; 
not possible to multiply two data frames using matrix multiplication. -->
  
  \vspace{60pt}

3. (2 points) Write an `R` code that sums the elements of the matrix `A` by row.
```{r}
A <- matrix(1:10000, 100, 100)
```

<!-- rowSums(A)  -->
  
  \vspace{60pt}

4. (2 points) Write down the result of the following `R` code.
```{r results = 'hide'}
x <- c(2, 2, 1, 1, 7, 1, 3, 2, 1, 7)
y <- c(rep("High", 5), rep("Low", 5))
table(x, y)
```

\vspace{60pt}

5. (2 points) Write an R code that, given a vector x with numerical values, 
   creates a vector y with entries equal to big (a character string) 
   if the corresponding entry of x is greater than 100, and small (a character string) otherwise. 
   Use the `ifelse()` function to do so.

   ```{r}

library(magrittr)
num_vec <- function(num_vec) {
  num_vec %<>% na.omit
  num_vec %>% sapply(function(x) {
    ifelse(x > 100, "Big", "Small")
  })
}

c(1, 200, 3, NA) %>% num_vec()
```

\vspace{60pt}


## Repetitive execution

(4 points) Write down the result of the following R code

```{r results = 'hide'}
x <- data.frame(X1 = 1:3, X2 = c(1.2, 3.0, 1.5), X2 = c(1.2, 3.0, 1.5), X3 = c("P", "Q", "R"))

sapply(1:dim(x)[1], function(i) {
  return(list(x[i, 2], paste(x[i, 1], x[i, 3], sep = "?")))
})
```

\vspace{60pt}

(4 points) In words, explain what the following code does

```{r, results="hide"}
rm(list = ls())
set.seed(1)
repetitions <- 1000
urn <- c(1:100)
trials_total <- 0

for (i in 1:repetitions) {
  num <- 0
  trials <- 0
  while (num != 55) {
    num <- sample(urn, 1, replace = FALSE)
    trials <- trials + 1
  }
  trials_total <- trials_total + trials
}

trials_total / repetitions
```

\vspace{60pt}

\newpage

## Functions

(4 points) Write a function called `descr_stats()` that takes a numeric vector and returns a vector of length 2 containing the sum of the numbers in the vector, the sample mean of the vector and the sample variance of the vector. Make sure that the function can return a number even if the vector contain NAs.

\vspace{120pt}



## Data Visualization 

1. (3 points) Describe what kind of a plot the following `R` code creates.

```{r fig.show = 'hide'}
library(ggplot2)
ggplot(data = mpg, aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(colour = class))
```



\vspace{120pt}

2. (5 points) Describe what kind of a plot the following `R` code creates.
```{r fig.show = 'hide'}
library(ggplot2)
ggplot(data = mpg, aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(colour = class)) +
  geom_text(data = mpg[mpg$hwy > 35, ], aes(x = displ, y = hwy, label = manufacturer))
```

<!-- It creates a scatterplot where each point is colored according to its class and where the manufacturers of the cars with hwy > 35 are displayed.  -->


\vspace{120pt}

